---
title: "Toronto's homeless death toll over the years shows the dangers of addiction's epidemic"
subtitle: "Data analysis of homeless deaths in Toronto from 2017 to 2023"
author: 
  - Shipeng Zhang
thanks: "Code and data are available at: https://github.com/zero616/Homeless_Death_Analysis."
date: 2024-01-22
date-format: long
abstract: "In Canada, the size and volatility of housing wealth has led many to believe that changes in housing prices have a significant impact on aggregate consumption. This study examines regional differences in housing price changes across Canada and their relationship to household spending patterns. Using forecasting models, we aim to estimate consumption levels in different regions based on NHPI data. However, limitations include potential inaccuracies in the NHPI data, assumptions about the relationship between house prices and consumption, and the model's ability to capture the full complexity of regional dynamics. Further research is needed to validate the model's predictions and explore other factors that influence consumer behavior."
format: pdf
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| include: false
#| warning: false
#| message: false
library("janitor")
library("knitr")
library("lubridate")
library("opendatatoronto")
library("tidyverse")
library(rstanarm)
```

# Introduction
In recent years, the Canadian real estate market has been the subject of extensive analysis and debate due to its significant impact on the economy and the well-being of its citizens. Canadian home prices have experienced significant volatility, raising concerns about affordability, financial stability and the broader impact on consumer consumption. Understanding the complex relationship between house prices and consumption is critical for policymakers, economists and stakeholders.

Canada's real estate market has experienced significant growth over the past few decades. Urban centers such as Toronto, Vancouver, and Montreal have become hot spots for real estate investment, driving up real estate values and reshaping the socioeconomic landscape. Historically low interest rates, demographic changes and foreign investment have fueled demand, leading to unprecedented price increases in some areas.

For many Canadians, housing is the most important component of household wealth. As property values rise, homeowners experience a wealth effect, feeling wealthier and more willing to spend money. This phenomenon has implications for consumption patterns, as increased housing wealth stimulates consumption on goods and services, thereby boosting economic growth. Conversely, falling house prices could lead to a contraction in consumer consumption, dampening economic activity.

Despite the wealth effect of rising home prices, affordability issues remain in the Canadian housing market. Soaring housing prices relative to income levels have made home ownership increasingly difficult for many Canadians, especially younger generations and immigrants. High housing costs not only strain household budgets but also limit discretionary consumption on other goods and services, potentially limiting overall consumption growth. At the same time, high rental costs force shop owners to raise rents, which in turn leads to rising prices.

Policy responses to housing market dynamics play a key role in shaping consumer behavior and broader economic outcomes. Measures aimed at cooling overheated housing markets, such as macroprudential regulation and tax policy, can influence housing affordability and household consumption decisions. Likewise, initiatives to boost the supply of affordable housing and enhance rental market stability are likely to reduce pressure on household budgets and support consumption growth.

Given the multifaceted nature of the relationship between house prices and consumption, this article aims to understand the dynamics in the Canadian context. By examining historical trends, empirical evidence, and the impact of emergencies, we seek to elucidate the interplay between housing market dynamics and consumer consumption and establish a rigorous link between the new home price index and the growth rate of Canadian consumer consumption. Analytical and comprehensive insights, this study strives to inform policymakers and stakeholders and contribute to a nuanced understanding of housing consumption relationships in Canada.


# Data {#sec-data}

This paper was produced using the R statistical programming language (R Core Team 2022).
here was used to reference file locations (Müller 2020). The data was examined and cleaned
using the packages janitor (Firke 2021), dplyr (Wickham et al. 2023), and tidyverse (Wickham
et al. 2019). Tables were made knitr (Xie 2023) and broom (0.7.11 2021), and formatted with
kableExtra (Zhu 2021). ggplot2 (Wickham 2016) was used to plot and scale the graphs.

## The Dataset
This dataset amalgamates information sourced from two key datasets available on the Canadian government's open platform: the "New Housing Price Index, Monthly" and the "Detailed Household Final Consumption Expenditure, Canada, Quarterly."

The New Housing Price Index (NHPI) is a monthly metric tracking variations in sales prices of newly constructed homes across time. It encompasses various housing categories, including detached houses, semi-detached houses, and townhouses. Additionally, the survey incorporates builders' assessments of the current market value of land, which are then independently indexed to create land value series. Similarly, the structural value of the properties is indexed and presented as a house series. Overall, the NHPI serves as a useful tool for gauging fluctuations within the Canadian real estate sector.

On the other hand, the Detailed Household Final Consumption Expenditure dataset provides a comprehensive summary of numerous expenditures incurred by Canadian residents for various purposes. It employs 2017 constant prices as a benchmark to eliminate the influence of inflation on growth rates.

Upon aggregating and refining the data to exclude irrelevant information, certain trends emerged. 
REF_DATE field denotes the specific quarter to which the data pertains, ensuring precision in analysis. 
Consumption figures represent the total consumption by Canadian residents during the quarter.km     
NHPI encapsulates Canada's national New Housing Price Index for the same period.

## Data Cleaning
```{r}
#| label: tbl-Sample-cost-raw
#| tbl-cap: Sample Table of Canada Consumtion change data (2000-2009)
#| warning: false
#| message: false
#| echo: false
#read file
cost_file_path <- "../../outputs/data/cleaned_data_cost.csv"

cost_raw_data <- read.csv(cost_file_path)


sample_data_1 <- head(cost_raw_data, 6)

# Print the table
knitr::kable(sample_data_1,'markdown',align = 'c')


```

```{r}
#| label: tbl-Sample-NHPI-raw
#| tbl-cap: Sample Table of Canada Consumtion and NHPI change (2000-2009)
#| warning: false
#| message: false
#| echo: false
#read file
NHPI_file_path <- "../../outputs/data/cleaned_data_NHPI.csv"
NHPI_raw_data <- read.csv(NHPI_file_path)
sample_data_2 <- head(NHPI_raw_data, 6)
# Print the table
knitr::kable(sample_data_2,'markdown',align = 'c')
```

```{r}
#| label: tbl-Sample
#| tbl-cap: Sample Table of Canada Consumtion and NHPI change (2000-2009)
#| warning: false
#| message: false
#| echo: false
#read file
file_path <- "../../outputs/data/analysis_data.csv"
analysis_data <- read.csv(file_path)
sample_data <- head(analysis_data, 6)
# Print the table
knitr::kable(sample_data,'markdown',col.names = c("Ref-Date", "Consumption (Million)", "Increasing Rate of Consumption", "NHPI", "Increasing Rate of Consumption"),align = 'c')


```

@tbl-Sample provides the total number of deaths summarized by age
group and cause of death in the data set, outlining the key variables
available for analysis: cause of death, age group and number of deaths.

```{r}
#| label: tbl-Consumption-increasing
#| tbl-cap: Sample Table for consumption data (2000-2009)
#| warning: false
#| message: false
#| echo: false

consumption_data <- analysis_data %>%
  select(-c(NHPI,Increasing_Rate_NHPI))
sample_consumption_data <- head(consumption_data, 6)
knitr::kable(sample_consumption_data,'markdown',col.names = c("Ref-Date", "Consumption (Million)", "Increasing Rate of Consumption"),align = 'c')
```

@tbl-Consumption-increasing provides the total number of deaths in the data set
summarized by year of death and cause of death, outlining the key
variables in the analysis of addiction harm to homeless people: cause of
death, time of death, and number of deaths.

```{r}

```
# Model
Null hypothesis: the relationship between the quarterly national NHPI growth rate and the growth rate of Canada's total final consumption is not linear
Alternative hypothesis: the relationship between the quarterly national NHPI growth rate and the growth rate of Canada's total final consumption is linear
Linear regression analysis was performed to test the hypothesis and determine whether there was a statistically significant relationship between the variables. A low p-value(typically less than a chosen significance level, often 0.05) for the linear regression coefficient will provide evidence against the null hypothesis, indicating
Changes in the NHPI growth rate do lead to meaningful increases in consumption by Canadian residents as a whole. On the other hand, if the p-value is high, it indicates that there is not enough evidence to reject the null hypothesis, and there may not be a significant linear relationship between the variables. Ultimately, the results of the analysis will inform whether the null hypothesis can or cannot be rejected.

## Model set-up
The polynomial regression model is defined as follows:
```{=tex}
\[ 
Y = \beta_0 + \beta_1X + \varepsilon 
\]

Where:
\begin{itemize}
  \item \( Y \) is the dependent variable (Increasing Rate of Consumption).
  \item \( X \) is the independent variable (Increasing Rate of NHPI).
  \item \( \beta_0 \) and \( \beta_1 \) are the coefficients of the polynomial terms.
  \item \( \varepsilon \) is the error term.
\end{itemize}
```

```{r}
#| label: tbl-model-summary
#| tbl-cap: Linear Regression Model Summary
#| warning: false
#| message: false
#| echo: false
  analysis_data <- read_csv("../../outputs/data/analysis_data.csv")
  linear_model <- lm(Increasing_Rate_Cost ~ Increasing_Rate_NHPI, data = analysis_data)
  # Summary of the model
  sample_model_summary<-summary(linear_model)$coefficients
  knitr::kable(sample_model_summary,'markdown',align = 'c')
```

@tbl-model-summary shows p-value is 0.0000001, then the linear model cannot be rejected.

## Model Justification
First observing the scatter plot of the data set, I believe that the relationship between the quarterly national NHPI growth rate and the growth rate of Canada's total final consumption is linear, then the linear regression model is appropriate.
Compared to Bayesian models, linear models are simpler and easier to interpret. It assumes a straight-line relationship between variables, which makes it easier to understand and communicate the results. In a linear model, the coefficient represents the change in the dependent variable when the predictor variable changes by one unit. This directly explains the impact of the NHPI growth rate on the growth rate of total final consumption in Canada. Linear regression is widely used and implemented in statistical software, making it easy to apply and interpret results. If the relationship between the NHPI growth rate and Canada's total final consumption growth rate is approximately linear, then a linear model can provide an adequate fit to the data. Analytical plots of predicted and actual values can help verify whether the linear model adequately captures the relationship.

## Model Prediction
```{r}
#| label: fig-model-prediction
#| fig-cap: Linear Regression Model Prediction
#| warning: false
#| message: false
#| echo: false
linear_model <- readRDS("../../models/model.rds")
# Get predicted values from the linear model
linear_model <- lm(Increasing_Rate_Cost ~ Increasing_Rate_NHPI, data = analysis_data)
  predicted_values <- predict(linear_model)
  analysis_data <- read_csv("../../outputs/data/analysis_data.csv")
  # Create a scatter plot of observed vs. predicted values
  ggplot(analysis_data, aes(x = Increasing_Rate_Cost, y = predicted_values)) +
    geom_point() +
    geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +  # Add diagonal line for reference
    labs(x = "Observed Consumption", y = "Predicted Consumption", title = "Observed vs. Predicted Consumption")
```
@fig-model-prediction shows the scatter plot compares the observed values of Consumption from the data set with the predicted values from the linear regression model. The dashed red line represents the line of perfect prediction (where observed equals predicted values). This plot allows us to visually assess how well the linear regression model fits the data. The points cluster closely around the dashed red line, indicating a strong fit between observed and predicted values. 

\newpage

# Results
```{r}
#| label: fig-model-plot
#| fig-cap: Scatter plot of Consumption vs. NHPI
#| warning: false
#| message: false
#| echo: false
 ggplot(analysis_data, aes(x = Increasing_Rate_NHPI, y = Increasing_Rate_Cost)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Add fitted line
    labs(x = "NHPI(Increasing Rate base 2000/1/1)", y = "Consumption(Increasing Rate base 2000/1/1)", title = "Scatter plot of Consumption vs. NHPI")
```

This research used a linear regression model to investigate the association between xxxxxxxx. As shown in Table 2, the estimated intercept produced by the model is 2.4367510, indicating that the expected value of the underlying consumption growth rate when the NHPI growth rate is zero is 2.4367510. The model also shows a statistically significant regression coefficient of 0.6191314, indicating that there is a significant linear relationship between the NHPI growth rate and the consumption growth rate. These findings indicate that as the NHPI growth rate increases, it is predicted that the consumption growth rate for the quarter will increase by approximately 0.6191314 percentage points, retaining all other variables. This shows that as the NHPI continues to grow, the final consumption of Canadian residents also continues to increase.
@fig-model-plot illustrates the relationship between NHPI growth rate and consumption growth rate, where the number of data points with NHPI growth rate on the x-axis and consumption growth rate on the y-axis of the scatterplot shows a clear upward slope. The fitted regression line represents the overall trend of the data, where the slope indicates that the consumption growth rate increases as the NHPI growth rate increases.


\newpage

# Discussion
```{r}
#| label: fig-consumption-plot
#| fig-cap: Consumption Over Time
#| warning: false
#| message: false
#| echo: false

fig_data<-analysis_data
fig_data$REF_DATE <- as.Date(paste(fig_data$REF_DATE, "-01", sep = ""), format = "%Y-%m-%d")
    ggplot(fig_data, aes(x = REF_DATE, y = Consumption)) +
    geom_bar(stat = "identity", fill = "skyblue", color = "black") +
    coord_cartesian(ylim = c(180000, max(fig_data$Consumption))) +
    labs(x = "Date", y = "Consumption")
```

```{r}
#| label: fig-NHPI-plot
#| fig-cap: NHPI Over Time
#| warning: false
#| message: false
#| echo: false

fig_data<-analysis_data
fig_data$REF_DATE <- as.Date(paste(fig_data$REF_DATE, "-01", sep = ""), format = "%Y-%m-%d")
    ggplot(fig_data, aes(x = REF_DATE, y = NHPI)) +
    geom_bar(stat = "identity", fill = "skyblue", color = "black") +
    coord_cartesian(ylim = c(55, max(fig_data$NHPI))) +
    labs(x = "Date", y = "NHPI")
```

## House Prices and Consumer consumption Trends

After analyzes historical trends in the New House Price Index (NHPI) and fluctuations in household consumer consumption. There is a significant correlation between the two.
The first point to focus on is the wealth effect, where homeowners typically experience an increase in wealth when house prices rise. Increased wealth can boost consumer confidence and discretionary consumption as homeowners feel more financially secure. Conversely, falling home prices may lead to less consumer consumption as homeowners feel less wealthy.
Second, fluctuations in the housing market, such as changes in housing affordability, mortgage rates, and housing supply and demand dynamics, may affect the NHPI and consumer consumption. For example, low mortgage rates may stimulate demand for housing, causing home prices to rise and increasing consumer consumption on related goods and services (e.g., home furnishings, renovations).
Consumer Sentiment: Consumer sentiment and perceptions of future economic prospects can influence consumption decisions. Positive sentiment may lead to increased consumer consumption, including housing-related expenditures, while negative sentiment may have the opposite effect. For older homeowners, rising house prices will stimulate an increase in consumer consumption. For young renters, rising house prices will lead to an increase in rents, and high rents will lead to a decrease in consumer consumption.

## Regional Differences

```{r}
#| label: fig-NHPI-toronto-plot
#| fig-cap: NHPI of Toronto Over Time
#| warning: false
#| message: false
#| echo: false

raw_data_NHPI_file_path <- "../../inputs/data/raw_data_NHPI.csv"

raw_data_NHPI <- read.csv(raw_data_NHPI_file_path)

cleaned_data_NHPI_Ontario <- raw_data_NHPI %>%
  filter(REF_DATE >="1999-12-01" & REF_DATE <="2009-10-01")
cleaned_data_NHPI_Ontario <- cleaned_data_NHPI_Ontario %>%
  filter(GEO == "Toronto, Ontario")
cleaned_data_NHPI_Ontario <- cleaned_data_NHPI_Ontario %>%
  filter(New.housing.price.indexes == "Total (house and land)")
cleaned_data_NHPI_Ontario <- cleaned_data_NHPI_Ontario %>%
  mutate(Increasing_Rate_NHPI = round(((VALUE / 57.2 - 1) * 100),2))
cleaned_data_NHPI_Ontario <- cleaned_data_NHPI_Ontario %>%
  select(-c(GEO,New.housing.price.indexes))
fig_data<-cleaned_data_NHPI_Ontario
fig_data$REF_DATE <- as.Date(paste(fig_data$REF_DATE, "-01", sep = ""), format = "%Y-%m-%d")
 # Create the line plot
ggplot(fig_data, aes(x = REF_DATE, y = VALUE)) +
  geom_line() +
  labs(x = "Date", y = "NHPI")
```
```{r}
#| label: fig-NHPI-Vancouver-plot
#| fig-cap: NHPI of Vancouver Over Time
#| warning: false
#| message: false
#| echo: false

raw_data_NHPI_file_path <- "../../inputs/data/raw_data_NHPI.csv"

raw_data_NHPI <- read.csv(raw_data_NHPI_file_path)

cleaned_data_NHPI_Vancouver <- raw_data_NHPI %>%
  filter(REF_DATE >="1999-12-01" & REF_DATE <="2009-10-01")
cleaned_data_NHPI_Vancouver <- cleaned_data_NHPI_Vancouver %>%
  filter(GEO == "Vancouver, British Columbia")
cleaned_data_NHPI_Vancouver <- cleaned_data_NHPI_Vancouver %>%
  filter(New.housing.price.indexes == "Total (house and land)")
cleaned_data_NHPI_Vancouver <- cleaned_data_NHPI_Vancouver %>%
  mutate(Increasing_Rate_NHPI = round(((VALUE / 57.2 - 1) * 100),2))
cleaned_data_NHPI_Vancouver <- cleaned_data_NHPI_Vancouver %>%
  select(-c(GEO,New.housing.price.indexes))
fig_data<-cleaned_data_NHPI_Vancouver
fig_data$REF_DATE <- as.Date(paste(fig_data$REF_DATE, "-01", sep = ""), format = "%Y-%m-%d")
# Create the line plot
ggplot(fig_data, aes(x = REF_DATE, y = VALUE)) +
  geom_line() +
  labs(x = "Date", y = "NHPI")
```

```{r}
#| label: fig-NHPI-Regina-plot
#| fig-cap: NHPI of Regina Over Time
#| warning: false
#| message: false
#| echo: false

raw_data_NHPI_file_path <- "../../inputs/data/raw_data_NHPI.csv"

raw_data_NHPI <- read.csv(raw_data_NHPI_file_path)

cleaned_data_NHPI_Regina <- raw_data_NHPI %>%
  filter(REF_DATE >="1999-12-01" & REF_DATE <="2009-10-01")
cleaned_data_NHPI_Regina <- cleaned_data_NHPI_Regina %>%
  filter(GEO == "Regina, Saskatchewan")
cleaned_data_NHPI_Regina <- cleaned_data_NHPI_Regina %>%
  filter(New.housing.price.indexes == "Total (house and land)")
cleaned_data_NHPI_Regina <- cleaned_data_NHPI_Regina %>%
  mutate(Increasing_Rate_NHPI = round(((VALUE / 57.2 - 1) * 100),2))
cleaned_data_NHPI_Regina <- cleaned_data_NHPI_Regina %>%
  select(-c(GEO,New.housing.price.indexes))
fig_data<-cleaned_data_NHPI_Regina
fig_data$REF_DATE <- as.Date(paste(fig_data$REF_DATE, "-01", sep = ""), format = "%Y-%m-%d")
# Create the line plot
ggplot(fig_data, aes(x = REF_DATE, y = VALUE)) +
  geom_line() +
  labs(x = "Date", y = "NHPI")
```

Studying how housing prices vary across different regions in Canada can offer valuable insights into local housing market dynamics and their impact on household spending habits. Analyzing historical housing price data across provinces and cities allows us to pinpoint areas with significant fluctuations in the National House Price Index (NHPI) over time. For instance, examining NHPI changes in Toronto, Vancouver, and Regina reveals distinct trends.
@fig-NHPI-toronto-plot shows the NHPI changes in Toronto. 
@fig-NHPI-Vancouver-plot shows the NHPI changes in Vancouver.
@fig-NHPI-Regina-plot shows the NHPI changes in Regina.
Comparing these trends, we observe that major cities with dense populations tend to experience more pronounced fluctuations in NHPI, while smaller towns with fewer inhabitants tend to have steadier curves. This variance may stem from the concentration of urban populations. Unfortunately, the Canadian government dataset lacks consumption data for provinces and cities. However, by leveraging the Vancouver we've developed and incorporating NHPI data from different regions, we can estimate consumption levels across various areas.

```{r}
#| label: fig-Cost-Toronto-plot
#| fig-cap: Predict Consumption of Toronto Over Time
#| warning: false
#| message: false
#| echo: false
# Step 2: Predict consumption using the polynomial regression model
linear_model <- lm(Increasing_Rate_Cost ~ Increasing_Rate_NHPI, data = analysis_data)
predicted_consumption <- predict(linear_model, newdata = cleaned_data_NHPI_Ontario)
# Create a dataframe for plotting
plot_data <- data.frame(Date = cleaned_data_NHPI_Ontario$REF_DATE, Predicted_Consumption = predicted_consumption)
fig_data <- plot_data
# Step 3: Create a linear plot using ggplot2
fig_data$Date <- as.Date(paste(fig_data$Date, "-01", sep = ""), format = "%Y-%m-%d")
ggplot(fig_data, aes(x = Date, y = Predicted_Consumption)) +
  geom_line() +
  labs(x = "Date", y = "Consumption")

```
```{r}
#| label: fig-Cost-Vancouver-plot
#| fig-cap: Predict Consumption of Vancouver Over Time
#| warning: false
#| message: false
#| echo: false
# Step 2: Predict consumption using the polynomial regression model
linear_model <- lm(Increasing_Rate_Cost ~ Increasing_Rate_NHPI, data = analysis_data)
predicted_consumption <- predict(linear_model, newdata = cleaned_data_NHPI_Vancouver)
# Create a dataframe for plotting
plot_data <- data.frame(Date = cleaned_data_NHPI_Vancouver$REF_DATE, Predicted_Consumption = predicted_consumption)
fig_data <- plot_data
# Step 3: Create a linear plot using ggplot2
fig_data$Date <- as.Date(paste(fig_data$Date, "-01", sep = ""), format = "%Y-%m-%d")
ggplot(fig_data, aes(x = Date, y = Predicted_Consumption)) +
  geom_line() +
  labs(x = "Date", y = "Consumption")

```

```{r}
#| label: fig-Cost-Regina-plot
#| fig-cap: Predict Consumption of Regina Over Time
#| warning: false
#| message: false
#| echo: false
# Step 2: Predict consumption using the polynomial regression model
linear_model <- lm(Increasing_Rate_Cost ~ Increasing_Rate_NHPI, data = analysis_data)
predicted_consumption <- predict(linear_model, newdata = cleaned_data_NHPI_Regina)
# Create a dataframe for plotting
plot_data <- data.frame(Date = cleaned_data_NHPI_Regina$REF_DATE, Predicted_Consumption = predicted_consumption)
fig_data <- plot_data
# Step 3: Create a linear plot using ggplot2
fig_data$Date <- as.Date(paste(fig_data$Date, "-01", sep = ""), format = "%Y-%m-%d")
ggplot(fig_data, aes(x = Date, y = Predicted_Consumption)) +
  geom_line() +
  labs(x = "Date", y = "Consumption")

```
@fig-Cost-toronto-plot shows the Predict Consumption changes in Toronto. 
@fig-Cost-Vancouver-plot x shows the Predict Consumption changes in Vancouver.
@fig-Cost-Regina-plot x shows the Predict Consumption changes in Regina.

## The Long-term Outlook

Assessing the long-term outlook for the Canadian housing market and household consumer consumption requires consideration of a range of key drivers and potential challenges that could influence future trends. Housing supply and demand dynamics include housing construction rates, land supply, zoning regulations, and housing affordability, which interact with demand-side factors to determine housing market equilibrium. Imbalances in housing supply and demand can lead to price volatility, affordability challenges and potential housing market corrections. Government policies and regulatory measures, such as housing subsidies, tax incentives, mortgage rules and land use regulations, can influence housing market dynamics and consumer consumption. Changes in government policy may introduce new incentives or restrictions that impact housing affordability and household finances. Developments in financial markets, including mortgage conditions, credit availability and mortgage rates, can affect housing market activity and consumer borrowing behavior. Changes in lending standards, interest rate policies and mortgage market liquidity may affect housing affordability and household debt levels. In the long term, technological advances and innovations in home construction, real estate transactions, and consumer finance can reshape housing markets and consumer consumption patterns. Digital platforms, fintech solutions and sustainable housing technologies may influence consumer preferences and market dynamics. Assessing the long-term outlook for the Canadian housing market and household consumer consumption requires consideration of the complex interplay of these factors and their potential impact on future trends. By monitoring key drivers and understanding potential challenges, policymakers, industry stakeholders and investors can better anticipate and adapt to changing market conditions and consumer preferences.
\newpage

## Next steps and weaknesses

This article lays the foundation for a more detailed study of the factors influencing final consumption of Canadian residents. Since there is no regional consumption data, it is impossible to confirm the impact of excessive fluctuations in housing prices on consumption data.
The next step would involve utilizing the model we've constructed to predict consumption levels in different regions based on NHPI data. This would entail feeding NHPI data from various regions into the model and generating predictions for consumption amounts in those regions. By doing so, we can gain insights into how housing price changes influence household spending patterns across Canada.

However, it's essential to acknowledge the weaknesses and limitations of this approach. One potential weakness is the assumption that NHPI changes directly correlate with changes in household spending patterns. While housing prices certainly impact consumer behavior, other factors such as income levels, employment rates, and economic policies also play significant roles.

Additionally, the model's accuracy may be affected by the quality and reliability of the NHPI data, as well as the assumptions and variables incorporated into the model itself. It's crucial to validate the model's predictions against real-world consumption data to assess its effectiveness and identify any potential biases or inaccuracies.

Furthermore, the model may not capture the full complexity of regional differences in housing markets and consumer behavior. Factors such as cultural preferences, demographic trends, and local economic conditions could influence consumption patterns in ways that aren't fully accounted for in the model.

Overall, while using NHPI data to predict consumption levels is a promising approach, it's important to interpret the results with caution and consider the broader context and potential limitations of the model.
## References
---
nocite: '@*'
---
```{r}
#| warning: false
#| message: false
#| echo: false
knitr::write_bib(c("base","knitr", "rmarkdown","janitor","lubridate","tidyverse","opendatatoronto","dplyr"), "references.bib")
new_entry <- "@misc{Opioid_overdoses_among_homeless_people_rose_dramatically_2023,
  title = {Opioid overdoses among homeless people rose dramatically: Ontario study},
  author = {},
  howpublished = {\\url{https://toronto.citynews.ca/2023/10/17/opioid-overdose-deaths-homeless-ontario/}},
  journal = {CityNews Toronto},
  year = {2023},
  month = {Oct},
  note = {Retrieved from \\url{https://toronto.citynews.ca/2023/10/17/opioid-overdose-deaths-homeless-ontario/}}
}"

# Append the new entry to the "references.bib" file
write(new_entry, "references.bib", append = TRUE)

```

